---
title             : "The Development of Color Terms in Shipibo-Konibo Children"
shorttitle        : "Color Terms in Shipibo-Konibo Children"

author: 
  - name          : "Danielle Kellier*"
    affiliation   : "1"
  - name          : "Martin Fortier*"
    affiliation   : "2"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Maria Fernández Flecha"
    affiliation   : "3"
  - name          : "Michael C. Frank"
    affiliation   : "4"

affiliation:
  - id            : "1"
    institution   : "University of Pennsylvania"
  - id            : "2"
    institution   : "PSL Research University"
  - id            : "2"
    institution   : "Pontificia Universidad Católica del Perú"
  - id            : "2"
    institution   : "Stanford University"

author_note: |
  * these authors contributed equally. 
  
  MCF was supported by a Jacobs Foundation Fellowship. We are especially grateful to Paul Kay for assistance throughout the study.

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, warning = FALSE, error = FALSE, 
                      message = FALSE, sanitize = TRUE, include = FALSE)
library("papaja")

library(tidyverse)
library(DT)
library(ggthemes)
library(cowplot)
library(viridis)
library(english)
library(lme4)
library(nnet)
library(here)

```

```{r analysis_preferences}
source(here("helpers/load_and_process.R"))

set.seed(42) # Seed for random number generation
predicted_range <- seq(6,10, 0.25)

theme_set(theme_bw() + 
            theme(strip.background = element_blank(), 
                  panel.grid = element_blank())) # nice theme with limited extras

```


TO BE PASTED FROM GOOGLE DOC


# Experiment 1

In our first experiment, our goal was to replicate and update the characterization of the adult SK color system given by the World Color Survey. We were further interested in the use of Spanish terms as language contact and multilingualism have increased in the years since the original World Color Survey work. 

```{r study1_participants}
study1_participants <- read_csv(
  here("data/Current_Data/naming_colors_participants.csv")) %>%
  mutate(age = as.numeric(as.character(edad)),
         `edad aprendizaje del espanol` = as.numeric(as.character(`edad aprendizaje del espanol`))) %>%
  summarise(mean_age = round(mean(age, na.rm = T), 1),
            median_age = round(median(age, na.rm = T), 1),
            sd_age = round(sd(age, na.rm = T), 1),
            min_age = min(age, na.rm = T),
            max_age = max(age, na.rm = T),
            mean_spanish = round(mean(`edad aprendizaje del espanol`, 
                                      na.rm = T), 1),
            median_spanish = round(median(`edad aprendizaje del espanol`, 
                                          na.rm = T), 1),
            sd_spanish = round(sd(`edad aprendizaje del espanol`, na.rm = T), 1),
            min_spanish = min(`edad aprendizaje del espanol`, na.rm = T),
            max_spanish = max(`edad aprendizaje del espanol`, na.rm = T),
            num_men = sum(género == 'masculino'),
            num_total = n())

study1_occupations <- read_csv(
  here("data/Current_Data/naming_colors_participants.csv")) %>%
  group_by(género, ocupación) %>%
  summarise (n = n()) %>%
  group_by(género) %>%
  mutate(freq_gender = 100*n / sum(n)) %>%
  ungroup() %>%
  mutate(freq_total = 100*n / sum(n))

```

## Methods 

### Participants

We recruited `r study1_participants$num_total` adult participants (`r study1_participants$num_men` men). Most of participants were from SK communities of the Middle Ucayali region (from Yarinacocha, San Francisco, and Nueva Betania), but some of them were from communities of the Lower (Paoyhan) and Upper (Puerto Belen) Ucayali. In Yarinacocha (a small town located in the vicinity of Pucallpa), participants were recruited in Bena Jema, a neighborhood where most of the inhabitants are SK. All the other places where participants were recruited were native community villages exclusively inhabited by SK people. Overall, the sample included both somewhat urbanized SK (Yarinacocha and San Francisco) as well as SK individuals who were still used to more traditional activities and regular contact with the surrounding rainforest (Nueva Betania, Paoyhan, and Puerto Belen).

The median self-reported age for participants was `r study1_participants$median_age` years with a range between `r study1_participants$min_age` to `r study1_participants$max_age` years of age (SD = `r study1_participants$sd_age`yo). Regarding occupations, `r round(filter(study1_occupations, género == 'femenino' & ocupación == 'ama de casa')$freq_gender)`% of the women were homemakers (`r round(filter(study1_occupations, género == 'femenino' & ocupación == 'ama de casa')$freq_total)`% overall) and another `r round(filter(study1_occupations, género == 'femenino' & ocupación == 'artesana')$freq_gender)`% were artisans (`r round(filter(study1_occupations, género == 'femenino' & ocupación == 'artesana')$freq_total)`%). `tools::toTitleCase(as.character(as.english(filter(study1_occupations, género == 'masculino' & ocupación == 'agricultor')$n)))` of the `r summarize(filter(study1_occupations, género == 'masculino'), n = sum(n))$n` men were horticulturalists (`r round(filter(study1_occupations, género == 'masculino' & ocupación == 'agricultor')$freq_gender)`%, `r round(filter(study1_occupations, género == 'masculino' & ocupación == 'agricultor')$freq_total)`% overall). `r tools::toTitleCase(as.character(as.english(filter(study1_occupations, género == 'femenino' & ocupación == 'estudiante')$n)))` women (`r round(filter(study1_occupations, género == 'femenino' & ocupación == 'estudiante')$freq_gender)`%, `r round(filter(study1_occupations, género == 'femenino' & ocupación == 'estudiante')$freq_total)`% overall) and `r as.character(as.english(filter(study1_occupations, género == 'masculino' & ocupación == 'estudiante')$n))` men (`r round(filter(study1_occupations, género == 'masculino' & ocupación == 'estudiante')$freq_gender)`%, `r round(filter(study1_occupations, género == 'masculino' & ocupación == 'estudiante')$freq_total)`% overall) identified as students. 

Although all adult participants spoke Shipibo-Konibo as a first language, all were bilingual to a substantial degree. All reported beginning to learn  Spanish before early adolescence (median = `r study1_participants$median_spanish`yo, SD = `r study1_participants$sd_spanish`y).

## Materials 

We used the 330 Munsell color chips as stimuli for the study. However, only 165 chips were used for each single participant (see below). These chips were exactly those used to collect data for the WCS. Individual color chips were 2 cm x 2.5 cm. 
<!-- The 330 hues of the set of chips can be visualized in Appendix 1.  -->


## Procedure

In order to make sure that the natural light intensity would not vary much between participants, the experiment took place indoors, near a window or a door. The study was conducted entirely in the SK language. 
 
Our procedure was similar to that used in the WCS [see @berlin2009, pp. 585--591].  Participants were seated in front of the experimenter and introduced to the whole procedure and the general goal of the study. Then the primary procedure involved presenting participants with a color chip and asking them: "What is the color of this chip?"^[The SK word for color that we used was the Spanish word *color*. In general, the SK language includes some castillanisms that are well-known by all speakers; color is one of them.] and recording their response or responses. 

One major difference between the WCS procedure and ours is that, in the WCS, the experimenter was expected to brief participants so that they would only provide basic color terms during the task (e.g., "blue" as opposed to "navy blue" or "sky-like"). We found it rather difficult to help participants understand in a few sentences what a basic color term was, however.^[Indeed, as Berlin & Kay (2009: 587-589) acknowledge, there is no straightforward necessary and sufficient criteria for the "basicness" of a color term.] Thus, we opted to let participants provide any term they wished. If they did not provide a basic color term, we would ask further questions to elicit a basic color term. For example, if, when presented with a red color chip, the participant provided the term "blood-like" (a non-basic color term), the experimenter would ask: "Do you know of any other word to refer to the color of this chip?" If the participant subsequently responded "dark red" (another non-basic color term), the experimenter would further ask: "How would you refer to this color with only one word?" Eventually, the participant would say "red" (a basic color term). 

For some chips, participants would provide a basic color term at once; but for others, they would first provide one or two non-basic terms before actually providing a basic term. When participants did not provide a basic color term after three trials (i.e., two follow-up questions), no further questions was asked, and the experimenter proceeded to the next chip. This method was more effortful and time-consuming than the WCS procedure, but it improved the fluency and the intuitiveness of the task for participants.

A second difference between our procedure and that of the WCS concerned the number of chips each participant was presented with. In the WCS, every participant was expected to provide color terms for each of the 330 chips of the set. As we were afraid that doing so would take too long and that participants would find the task tedious, we decided that the set of chips would be split in two (even and uneven numbers) and that every participant would be randomly ascribed to one of the two subsets. As a result, each participant was presented with only 165 chips.

## Results and Discussion

```{r study1_commonterms}

study1_profusion <- naming_data %>%
  group_by(subj, color_cat) %>%
  summarise(n = n()) %>%
  group_by(color_cat) %>%
  spread(subj, n, fill = 0) %>%
  gather(key = 'subj', value = 'n', -color_cat) %>%
  mutate(prop = 100*n/165) %>%
  summarise(`% of Subjects Who Used the Term (Current)` = 100*sum(prop > 0)/n(), 
            `Mean % of Chips in Set Labeled (Current)` = mean(prop),
            `SD % of Chips in Set Labeled (Current)` = sd(prop))

curr_naming_list <- as.character(na.omit(filter(study1_profusion, `% of Subjects Who Used the Term (Current)` > 50 & !is.na(color_cat))$`color_cat`))

study1_switchperc <- naming_data %>%
  mutate(spanish_response = ifelse(color_cat %in% spanish_terms,
                                   1, 0)) %>%
  mutate(overall_spanish = paste0(round(100*mean(spanish_response), 0),"%")) %>%
  group_by(subj) %>%
  summarise(overall_spanish = first(overall_spanish),
            atleast_1 = ifelse(sum(spanish_response) > 0, 1, 0),
            by_subject = mean(spanish_response)) %>% 
  ungroup() %>%
  summarise(overall_spanish = first(overall_spanish),
            atleast_1 = paste0(round(100*mean(atleast_1), 0),"%"),
            mean_subj_spanish = paste0(round(100*mean(by_subject)),"%"),
            sd_subj_spanish = paste0(round(100*sd(by_subject)),"%"),
            max_subj_spanish = paste0(round(100*max(by_subject)),"%"),
            min_subj_spanish = paste0(round(100*min(by_subject)),"%"))

study1_spanishbychip <- naming_data %>%
  mutate(spanish_response = ifelse(color_cat %in% spanish_terms,
                                   1, 0)) %>%
  select(subj, chip_id, spanish_response) %>%
  mutate(set = ifelse((chip_id %% 2) == 0, 'even', 'odd')) %>%
  split(.$set) %>%
  map_df(function(x) {
    x %>%
      group_by(chip_id) %>%
      summarise(perc = mean(spanish_response))
  }) %>%
  left_join(color_chip_data %>% select("#cnum", "hex"),
            by = c("chip_id" = "#cnum")) %>%
  arrange(desc(perc)) %>%
  rename(`Chip ID` = chip_id, `% of Subjects` = perc, `Hex Code` = hex)

study1_highestspanish <- datatable(study1_spanishbychip, rownames = FALSE) %>%
  formatPercentage("% of Subjects", 0) %>%
  formatStyle(columns = "Hex Code",
              background = styleEqual(
                study1_spanishbychip$`Hex Code`, study1_spanishbychip$`Hex Code`))

```

```{r study1_moredetail}

study1_consensus <- naming_data %>%
  select(subj, chip_id, color_cat) %>%
  mutate(set = ifelse((chip_id %% 2) == 0, 'even', 'odd')) %>%
  split(.$set) %>%
  map_df(function(x) {
    x %>%
      group_by(chip_id, color_cat) %>%
      summarise(n = n()) %>%
      group_by(chip_id) %>%
      mutate(perc = 100*n/sum(n)) %>%
      select(-n)
  }) %>%
  arrange(chip_id) %>%
  rename(`Chip ID` = chip_id, `Color Term` = color_cat, `% of Subjects` = perc) 


```

```{r adultfigure, fig.cap = "(A and B) Plots of the modal term given for a particular chip. Color coordinates were represented in 2-D Munsell space. Modal responses were given by SK adults during (A) the original World Color Survey and during (B) our Experiment 1. (C) Heat map of prevalence of Spanish-language responses during Experiment 1. Legends for all three subplots located in the bottom-right quadrant."}

study1_consensus_plot <- ggplot(study1_consensus %>%
                                 group_by(`Chip ID`) %>%
                                 filter(`% of Subjects` == max(`% of Subjects`)) %>%
                                 left_join(color_chip_data, by = c("Chip ID" = "#cnum")), 
                               aes(x = H, y = V, fill = `Color Term`)) + 
  geom_tile() + 
  scale_size(range = c(0, 2.0)) + 
  scale_fill_manual(name = "Color Term",values = graph_colors) +
  ggtitle("Experiment 1") +
  guides(fill=guide_legend(ncol=2,byrow=TRUE)) +
  theme(plot.title = element_text(hjust = 0.5))

kay_consensus <- kay_terms %>%
  mutate(Term = ifelse(Term %in% c('Huiso', 'Wiso'), 'Wiso', as.character(Term))) %>%
  group_by(`WCS Chip Number`, Term) %>%
  summarise(n = n()) %>%
  group_by(`WCS Chip Number`) %>%
  mutate(`% of Subjects` = 100*n/sum(n)) %>%
  select(-n) %>%
  group_by(`WCS Chip Number`) %>%
  filter(`% of Subjects` == max(`% of Subjects`)) %>%
  dplyr::rename(`Color Term` = Term, `Chip ID` = `WCS Chip Number`) %>%
  left_join(color_chip_data, by = c("Chip ID" = "#cnum"))


kay_consensus_plot <- ggplot(kay_consensus, 
                             aes(x = H, y = V, fill = `Color Term`)) + 
  geom_tile() + 
  scale_size(range = c(0, 2.0)) + 
  scale_fill_manual(name = "Color Term",values = graph_colors) +
  ggtitle("WCS Data") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

study1_spanish_plot <- ggplot(study1_spanishbychip %>%
                               mutate(`% of Subjects` = 100*`% of Subjects`) %>%
                               rename(`Spanish Term` = `% of Subjects`) %>%
                               left_join(select(color_chip_data, `#cnum`, V, H), 
                                         by = c("Chip ID" = "#cnum")), 
                             aes(x = H, y = V)) + 
  geom_tile(aes(fill = `Spanish Term`)) + 
  scale_fill_viridis(option="magma") +
  labs(fill = "Spanish\nUse (%)")

adult_figurelegends <- plot_grid(
  get_legend(study1_consensus_plot),
  NULL,
  get_legend(study1_spanish_plot),
  labels = c("A and B", "", "C"),
  ncol = 3) + 
  theme(plot.margin = unit(c(0.5,0.5,0.5,1), "in")) 

adult_figure <- plot_grid(
  study1_consensus_plot + theme(legend.position = "none"), 
  kay_consensus_plot + theme(legend.position = "none"), 
  study1_spanish_plot + theme(legend.position = "none"), 
  adult_figurelegends, 
  labels = c("A", "B", "C"), ncol = 2
  )

print(adult_figure)

```


Broadly speaking, our results were quite similar to the WCS findings. Figure 1 shows a comparison between our data (Panel A) and the WCS (panel B). The basic level colors in our data were quite similar, as well. All participants described at least 1 chip with the following set of color terms: light/white ("joxo"), dark/black ("wiso"), yellow ("panshin"), red ("joshin"), and green/blue ("yankon"). Most (`r paste0(round(filter(study1_profusion, color_cat == "Manxan")$'% of Subjects Who Used the Term (Current)'), '%')`) participants also used described at least 1 chip as faded or "manxan", referring to a chip's saturation. In terms of overall popularity, participants on average described `r paste0(round(filter(study1_profusion, color_cat == "Yankon")$'Mean % of Chips in Set Labeled (Current)'), '%')` of chips as "yankon" (*SD* = `r paste0(round(filter(study1_profusion, color_cat == "Yankon")$'SD % of Chips in Set Labeled (Current)'), '%')`) followed by "joshin" (*M* = `r paste0(round(filter(study1_profusion, color_cat == "Joshin")$'Mean % of Chips in Set Labeled (Current)'), '%')`, *SD* = `r paste0(round(filter(study1_profusion, color_cat == "Joshin")$'SD % of Chips in Set Labeled (Current)'), '%')`), "joxo" (`r paste0(round(filter(study1_profusion, color_cat == "Joxo")$'Mean % of Chips in Set Labeled (Current)'), '%')`, `r paste0(round(filter(study1_profusion, color_cat == "Joxo")$'SD % of Chips in Set Labeled (Current)'), '%')`), "panshin" (`r paste0(round(filter(study1_profusion, color_cat == "Panshin")$'Mean % of Chips in Set Labeled (Current)'), '%')`, `r paste0(round(filter(study1_profusion, color_cat == "Panshin")$'SD % of Chips in Set Labeled (Current)'), '%')`), "manxan" (`r paste0(round(filter(study1_profusion, color_cat == "Manxan")$'Mean % of Chips in Set Labeled (Current)'), '%')`, `r paste0(round(filter(study1_profusion, color_cat == "Manxan")$'SD % of Chips in Set Labeled (Current)'), '%')`), and "wiso" (`r paste0(round(filter(study1_profusion, color_cat == "Wiso")$'Mean % of Chips in Set Labeled (Current)'), '%')`, `r paste0(round(filter(study1_profusion, color_cat == "Wiso")$'SD % of Chips in Set Labeled (Current)'), '%')`).

One departure from the Berlin-Kay data was that `r study1_switchperc$atleast_1` of adults described at least 1 chip using a Spanish-language color term, accounting for `r study1_switchperc$overall_spanish` of all responses (Figure 1, Panel C). In particular, Spanish use reached as high as `r round(100*max(study1_spanishbychip$'% of Subjects'))`% when participants were asked to label chips that English speakers would consider to be orange. However, there was a high amount of variability in Spanish use between subjects (*M* = `r study1_switchperc$mean_subj_spanish`, *SD* = `r study1_switchperc$sd_subj_spanish`) with some subjects never responding in Spanish. One responded in Spanish for `r study1_switchperc$min_subj_spanish` of all trials despite all sessions being conducted entirely in the Shipibo-Konibo language. While we can only speculate as to this participant's motivations, it seems likely that they were more familiar with the Spanish vocabulary or viewed it as more precise. 

<!--<br>

`r study1_highestspanish` 
<br>-->


```{r study1_termnature}
study1_nature <- naming_data %>%
  mutate(spanish_response = ifelse(color_cat %in% spanish_terms, 1, 0)) %>%
  group_by(subj, spanish_response, nature) %>%
  summarise(n = n()) %>% ungroup() %>%
  mutate(nature = ifelse(is.na(nature), "Other", nature)) %>%
  complete(subj, spanish_response, nature, fill = list(n = 0)) %>%
  group_by(subj) %>%
  mutate(perc = 100*n/sum(n)) %>%
  group_by(spanish_response, nature) %>%
  summarise(mean_prop = paste0(round(mean(perc)), '%'),
            sd_prop = paste0(round(sd(perc)), '%'))
  
```

Participants on average described `r filter(study1_nature, nature == "BCT" & spanish_response == 0)$mean_prop` of chips using a SK-language basic color term like "yankon" (*SD* = `r filter(study1_nature, nature == "BCT" & spanish_response == 0)$sd_prop`). Some participants described chips using SK-language ad hoc color terms, such as "nai" or *sky* for blue chips (*M* = `r filter(study1_nature, nature == "AHCT" & spanish_response == 0)$mean_prop`, *SD* = `r filter(study1_nature, nature == "AHCT" & spanish_response == 0)$sd_prop`), or ad hoc terms referring to saturation or luminosity of a chip, such as "manxan" (*M* = `r filter(study1_nature, nature == "AHCT - SLT" & spanish_response == 0)$mean_prop`, *SD* = `r filter(study1_nature, nature == "AHCT - SLT" & spanish_response == 0)$sd_prop`). Virtually all instances where a participant responded in Spanish involved a Spanish basic color term such as "rojo" (*M* = `r filter(study1_nature, nature == "BCT" & spanish_response == 1)$mean_prop`, *SD* = `r filter(study1_nature, nature == "BCT" & spanish_response == 1)$sd_prop`). In other words, participants typically only responded in Spanish to label chips into basic categories; they relied on Shipibo-Konibo for other descriptors.


```{r, study1_genderdifferences, include = FALSE}
study1_genderbychip <- naming_data %>%
  group_by(género) %>%
  mutate(gender_count = n_distinct(subj)) %>%
  group_by(chip_id, género, color_cat, gender_count) %>%
  summarise(n = n()) %>%
  mutate(prop = n/gender_count) %>%
  select(-gender_count, -n) %>% ungroup() %>%
  complete(chip_id, género, color_cat, fill = list(prop = 0))

study1_genderbyterm <- naming_data %>%
  group_by(subj, género, color_cat) %>%
  summarise(prop = n()/165) %>% ungroup() %>%
  complete(nesting(subj, género), color_cat, fill = list(prop = 0))

study1_ttest_gendertermprofusion <- with(
  study1_genderbyterm %>% 
   group_by(color_cat, género) %>%
   summarise(prop = mean(prop)) %>%
   spread(género, prop), t.test(femenino, masculino, paired = TRUE))

study1_ttest_gendertermappearance <- with(
  study1_genderbyterm %>% 
    mutate(used = ifelse(prop > 0, 1, 0)) %>%
    group_by(color_cat, género) %>%
    summarise(prop = mean(used)) %>%
    spread(género, prop), t.test(femenino, masculino))

```

<!-- We found no significant gender differences in the overall spread of color term usage across the set (*t* = `r study1_ttest_gendertermprofusion$statistic`, *P* = `r study1_ttest_gendertermprofusion$p.value`) or in the percentage of subjects who used a term at least once throughout the task (*t* = `r study1_ttest_gendertermappearance$statistic`, *P* = `r study1_ttest_gendertermappearance$p.value`). -->
<!-- MCF: There are some issues with these stats - you really want to test for distributional differences, but it's complicated. -->

```{r study1_corrected_entropy}
study1_entropy <- naming_data %>%
  group_by(chip_id, color_cat) %>% 
  select(chip_id, color_cat) %>%
  count %>%
  group_by(chip_id) %>%
  mutate(p = n / sum(n)) %>%
  summarise(entropy = - sum(p * log(p))) %>%
  rename(`Chip ID` = chip_id, 
         Entropy = entropy)
```

```{r, eval=FALSE}
study1_naturebyentropy <-naming_data %>%
  mutate(response_language = case_when(
    color_cat %in% spanish_terms ~ "spanish",
    color_cat %in% shipibo_terms ~ "shipibo"
  )) %>%
  select(subj, chip_id, color_cat, response_language, nature) %>%
  mutate(shipibo_BCT = ifelse(nature == "BCT" & response_language == "shipibo", 1, 0),
         shipibo_AHCT = ifelse(nature == "AHCT" & response_language == "shipibo", 1, 0),
         spanish_BCT = ifelse(nature == "BCT" & response_language == "spanish", 1, 0),
         spanish_AHCT = ifelse(nature == "AHCT" & response_language == "spanish", 1, 0)) %>%
  mutate_at(vars(shipibo_BCT:spanish_AHCT), funs(ifelse(is.na(.), 0, .))) %>%
  group_by(chip_id) %>%
  summarise_at(vars(ends_with("CT", ignore.case = FALSE)), mean, na.rm = TRUE) %>%
  right_join(study1_entropy, by = c("chip_id" = "Chip ID")) %>%
  select(`Chip ID` = chip_id, Entropy, ends_with("CT", ignore.case = FALSE))

## A little circular for a correlation?
study1_naturebyentropy_test <- cor.test(~ shipibo_BCT + Entropy, data = study1_naturebyentropy)

study1_naturebyentropy_plot <- ggplot(study1_naturebyentropy %>% 
         gather(key = "measure", value = "prop", ends_with("CT")) %>%
         separate(measure, c("language", "nature"), sep = "_") %>%
         mutate(language = tools::toTitleCase(language),
                nature = case_when(
                  nature == "BCT" ~ "Basic color term",
                  nature == "AHCT" ~ "Ad-hoc color term"),
                language = relevel(factor(language), ref = "Shipibo"),
                nature = relevel(factor(nature), ref = "Basic color term")),
       aes(x = Entropy, y = prop, group = interaction(language, nature), colour = language, shape = nature)) +
  geom_smooth(method = "loess", aes(fill = language, linetype = nature), alpha = .2) +
  labs(x = "Labelling Entropy of Chip", y = "Proportion of Responses", fill = "Language", color = "Language", linetype = "Nature of Response") +
  scale_fill_viridis(discrete = TRUE, option = "D") +
  scale_color_viridis(discrete = TRUE, option = "D") +
  coord_cartesian(ylim = c(0,1))

print(study1_naturebyentropy_plot)

```

Given these data, we next moved on to exploring the development of SK color vocabulary in childhood. Experiment 2 tests production and comprehension of SK color terms using SK-prototypical color chips; Experiment 3 tests children in Spanish using Spanish-prototypical chips. 

# Experiment 2

In Experiment 2, we tested children on their production and comprehension skills with a set of chips representing the prototypical colors for common SK color terms. 

```{r study2_participants}

study2_data <- feather::read_feather("../intermediates/processed_data.csv") %>%
  filter(language == "Shipibo", task == "Production") %>%
  mutate(spanish_response = ifelse(response %in% spanish_terms, 1, 0)) 

study2_participants <- study2_data %>%
  group_by(subj) %>%
  filter(row_number() == 1) %>% ungroup() %>%
  summarise(min_age = min(age_interval),
            max_age = max(age_interval),
            num_56 = sum(age_interval %in% 5:6),
            num_78 = sum(age_interval %in% 7:8),
            num_910 = sum(age_interval %in% 9:10),
            num_1112 = sum(age_interval %in% 11:12),
            num_benajema = sum(location == 'bena jema'),
            num_bilingue = sum(location == 'bilingue'),
            num_ccnn = sum(location == 'c.c.n.n. (bahuanisho)'),
            num_boys = sum(gender == "masculino"),
            num_total = n())

study2_agesex <- study2_data %>%
  group_by(subj) %>%
  filter(row_number() == 1) %>% group_by(age_interval) %>%
  summarise(num_total = n(),
            num_boys = sum(gender == 'masculino'),
            mean_age = round(mean(age), 1),
            sd_age = round(sd(age), 2)) %>% ungroup() %>%
  mutate(freq_total = round(100*num_total/sum(num_total)),
         freq_boys = round(100*num_boys/num_total)) %>%
  mutate(`Age Group` = as.character(age_interval),
         `N` = paste0(num_total, " (", freq_total, "%)"),
         `Male` = paste0(num_boys, " (", freq_boys, "%)")) %>%
  select(`Age Group`, `N`, `Male`)

study2_table1 <- apa_table(study2_agesex, caption = "Demographics of participants in Experiment 2.")

```

## Methods

### Participants

The Pontificia Universidad Católica del Perú’s Institutional Review Board approved our study protocol. We recruited `r study2_participants$num_total` `r study2_participants$min_age`- to `r study2_participants$max_age`-year-old children (`r study2_participants$num_boys` boys). Table 1 shows the distribution of ages and genders. `r tools::toTitleCase(as.character(as.english(with(study2_participants, num_benajema + num_bilingue))))` children were recruited from neighborhoods in Yarinacocha, in the Pucallpa region of Peru, as well as in `r study2_participants$num_ccnn` children from Bawanisho, a native community settled along the Ucayali River, south of Pucallpa. Children were recruited either through their parents or through local schools. When recruited at school, consent for participation was collected from both the teachers and the parents; otherwise, only consent from the parents was collected.  

### Materials

Based on findings of Experiment 1, we selected out 8 color chips that were prototypical instances of prominent SK color terms. These color chips were blue (WCS n°1), green (WCS n°234), red (WCS n°245), white (WCS n°274), yellow (WCS n°297), black (WCS n°312), greeny-yellow (WCS n°320), and purple (WCS n°325). These color chips were exactly the same as those used in Experiment 1; the only difference was that while 330 chips were used in Experiment 1, only 8 were used in Experiment 2.

### Procedure

The production and comprehension tasks were both conducted in SK. In both tasks, children were seated in front of the experimenter. A table on which the color chips were display stood between them. The production task was always performed before the comprehension task.

**Production task.** The procedure was very similar to that of Experiment 1. Children were first introduced to the whole procedure and the general goal of the study. It was specified that they would be expected to provide color terms in SK (and not in Spanish). Children were then asked: "What is the color of this chip?". As with adults, we used follow-up questions to elicit basic color terms when the terms children initially provided were not basic. When children provided Spanish color terms, the experimenter would write down their response but further ask: "What is the name of this color in SK?" When children replied "I don’t know" to this prompt, the experimenter would not ask further questions and would move forward to the next color chip. As a result, responses of some children include only non-basic SK color terms or Spanish color terms. In total, we collected production data for 8 color chips. For each chip, the data include either one response (when children provided a SK basic color term in the first trial) or two or three responses (when children’s initial responses were either non-basic and/or in Spanish).

**Comprehension task.** The 8 color chips of the production task were simultaneously displayed in front of the children. The experimenter would then ask: "Can you give me the [color] chip?" In total, the comprehension of 9 SK color terms was tested. The choice of these terms was based on the findings of Experiment 1. Not all of them were basic, but all of them stood out as being prominent in the SK color system. The 9 terms used as prompts included: yankon ("grue"), joshin ("red"), panshin ("yellow"), joxo ("white), wiso ("black"), nai ("blue"), and barin poi ("greeny-yellow"). In addition, as Experiment 1 revealed that two non-basic terms are widely used to refer to green and purple, two words were used to test comprehension of each of these two colors: pei/xo ("green") and ami/pua ("purple").

Further, Experiment 1 showed that for some of these color terms, only one response was accurate, while for others, several responses were equally correct. For example, only one chip could be picked up as an instance of wiso. By contrast, four chips could be considered to be instances of yankon (blue, green, greeny-yellow, and, to a lesser extent, purple); two chips for joshin (red, and, to a lesser extent, purple); and two as well for pei/xo (green, and, to a lesser extent, green-yellow). Accuracy was coded based on the results derived from Experiment 1: if at least 15% of participants in Experiment 1 associated a chip with a particular label, we considered a trial to be correct if the child made the same pairing.

When the experimenter asked children to pick up a color that was instantiated by several chips, we followed the following procedure. The experimenter would ask: "Can you give me the [color] chip?" Children would then pick up a chip. The response would be registered and the chip be taken out of the table. As a result, only 7 chips would be remaining on the table. The experimenter would subsequently ask: "Can you give me another [color] chip?". Children would then pick up a new chip. The response would be registered and the chip be taken out of the table. The experimenter would then ask the same question again until a total of as many times as there were correct instances. Thus, for example, for yankon four chips would be elicited, while for joshin, two chips would be elicited. 


## Results and Discussion

```{r childfigure, fig.cap = "A comparison of children's performance during the production task in Studies 2 and 3. Solid or dotted lines represent overall performance by age for a particular chip. Solid lines show whether the child gave a correct answer in the language indicated in that column; dotted lines show if they gave a response that was correct in either language. Line colors are representative of the chip’s color coordinates. Values in the lower-right corners of each subplot display the entropy (uncertainty) values calculated from adult responses given during Experiment 1."}

all_chip_sets <- full_join(shipibo_chip_set, 
                           spanish_chip_set) %>%
  left_join(color_chip_data, 
            by = c("chip_id" = "#cnum"))

chip_id_colors <- setNames(all_chip_sets$hex, 
                           all_chip_sets$chip_id)

study23_firstprod <- feather::read_feather("../intermediates/processed_data.csv") %>% 
  filter(task == 'Production' & 
           response_order %in% c('first','only') &
           spanish %in% c("Blanco", "Negro", 
                          "Verde", "Amarillo", "Rojo",
                          "Morado")) %>%
  group_by(language, age, prompt) %>%
  rename(`Correct (Same Language)` = correct, 
         `Correct (Either Language)` = correct_either) %>%
  gather(key = "criteria", value = "score", 
         `Correct (Same Language)`, `Correct (Either Language)`) %>%
  group_by(language, prompt, age_interval, criteria) %>%
  left_join(study1_entropy %>% 
              mutate(`Chip ID` = as.character(`Chip ID`),
                     Entropy = sprintf("%.2f", round(Entropy,2))), 
            by = c("prompt" = "Chip ID"))

ggplot(study23_firstprod,
                       aes(x = age_interval, y = score, colour = prompt,
                           group = criteria, linetype = criteria)) +
  geom_jitter(data = filter(study23_firstprod, criteria == "Correct (Same Language)"),
    alpha = .25, width = .3, height = .1, col = "gray") + 
  geom_smooth(se = FALSE, method = "loess", span = 1) +
  facet_grid(as.numeric(as.character(prompt)) ~ language) + 
  scale_y_continuous(limits = c(-0.1,1.1), breaks = seq(0,1,0.5), labels = scales::percent) + 
  xlim(c(5,11.5)) +
  scale_color_manual(values = chip_id_colors) +
  scale_linetype_manual(values = c("Correct (Same Language)" = "solid", 
                                   "Correct (Either Language)" = "dotted"), 
                        guide = FALSE) +
  labs(x = "Age", y = "% of Correct Responses", 
       linetype = "Measure", colour = "Chip ID") +
  guides(label = "none") + 
  geom_label(aes(x = Inf, y = -Inf, hjust = 1.1, vjust = -0.3, label = Entropy),
             show.legend = FALSE, colour = "black")

```

```{r}
study23_firstprod$age.years.c <- scale(study23_firstprod$age, scale = FALSE)
firstprod_same <- glmer(score ~ age.years.c  + (age.years.c | prompt) + (1 | subj), 
      data = filter(study23_firstprod, criteria == "Correct (Same Language)"), 
      family = "binomial")
firstprod_either <- glmer(score ~ age.years.c  + (age.years.c | prompt) + (1 | subj), 
      data = filter(study23_firstprod, criteria == "Correct (Either Language)"), 
      family = "binomial")
```

**Production.** Children's production accuracy increased substantially across nearly all color chips in the age range that we tested. Figure 2, left panels shows the accuracy of children's first production, both in SK (solid line) and in either language (dashed line). To quantify these developmental trends, we fit two generalized linear mixed effects models, one for the accuracy of SK production and one for the accuracy of production in either language. Both of these predicted accuracy as a function of the child's age, and included random intercepts for color chip and for participant, as well as a random slope of age by color chip. Age was a significant predictor in both models: $\beta = `r round(summary(firstprod_same)$coef["age.years.c", "Estimate"], digits = 2)`$, SE = `r round(summary(firstprod_same)$coef["age.years.c", "Std. Error"], digits = 2)`, $p = `r round(summary(firstprod_same)$coef["age.years.c", "Pr(>|z|)"], digits = 3)`$ and $\beta = `r round(summary(firstprod_either)$coef["age.years.c", "Estimate"], digits = 2)`$, SE = `r round(summary(firstprod_either)$coef["age.years.c", "Std. Error"], digits = 2)`, $p < .0001$.


```{r study2_accuracyspanishentropy}

study2_overallscore <- study2_data %>%
  group_by(subj, age) %>%
  summarise(correct = mean(correct, na.rm = T))

study2_overallscore_test <- cor.test(~ correct + age, data = study2_overallscore)

study2_overallswitch <- study2_data %>%
  group_by(subj, age) %>%
  summarise(spanish_response = mean(spanish_response))

study2_overallswitch_test <- cor.test(~ spanish_response + age, data = study2_overallswitch)

study2_switchperc <- study2_data %>%
  summarise(spanish_response = paste0(round(100*mean(spanish_response)), "%"))

study2_responsebyentropy <- study2_data %>%
  filter(spanish %in% c("Blanco", "Negro", "Verde", "Amarillo", "Rojo") & 
           (task == "Comprehension" | response_order %in% c("first", "only"))) %>%
  left_join(study1_entropy %>% mutate(`Chip ID` = as.character(`Chip ID`)),
            by = c("prompt" = "Chip ID"))

study2_responsebyentropy$age.years.c <- scale(study2_responsebyentropy$age, scale = FALSE)[,1]

study2_SwitchEnt_model <- summary(lme4::glmer(
  spanish_response ~ age.years.c * Entropy + (1|subj) + (1 | prompt),
  data=study2_responsebyentropy, family = "binomial"))
```

Over a quarter (`r study2_switchperc$spanish_response`) of all responses were given in Spanish, and the distribution of Spanish responses was non-random. Children tended to respond in Spanish when presented with a chip with low naming consensus among adult participants in Experiment. We computed the naming entropy for each chip by computing the probabilities for each chip $c$ to be named with a particular label $l$ ($p(l \mid c)$) and then taking $H(c) = - \sum{p(l\mid c) \log[p(l \mid c)]}$ (see inset entropy values by chip in Figure 2). 

To assess the hypothesis that naming entropy in adults was related to Spanish use in children, we fit a mixed effects model predicting Spanish responses as a function of age, entropy of the chip's naming distribution for adults, and their interaction. We included random intercepts for color chip and for participant, but our model did not converge with a random slope term and so we pruned this term following our lab's standard operating procedure. We found a reliable effect of entropy ($\beta = `r round(summary(study2_SwitchEnt_model)$coef["Entropy", "Estimate"], digits = 2)`$, SE = `r round(summary(study2_SwitchEnt_model)$coef["Entropy", "Std. Error"], digits = 2)`, $p = `r round(summary(study2_SwitchEnt_model)$coef["Entropy", "Pr(>|z|)"], digits = 3)`$) and an interaction between age and entropy ($\beta = `r round(summary(study2_SwitchEnt_model)$coef["age.years.c:Entropy", "Estimate"], digits = 2)`$, SE = `r round(summary(study2_SwitchEnt_model)$coef["age.years.c:Entropy", "Std. Error"], digits = 2)`, $p = `r round(summary(study2_SwitchEnt_model)$coef["age.years.c:Entropy", "Pr(>|z|)"], digits = 3)`$), suggesting that adults' uncertainty regarding naming was related to children's likelihood of producing Spanish labels. 



<!-- **Overextensions.** -->
```{r study2_overextensions}

study2_adjacentorswitch_whole <- study2_data %>%
  group_by(subj, age, gender) %>%
  filter(response_order %in% c("first", "only")) %>%
  mutate(correct_nearby = correct + ifelse(is.na(wrong_adjacent), 0, wrong_adjacent),
         correct_all = ifelse(correct + correct_either + correct_nearby > 0, 1, 0)) %>%
  select(subj, age, gender, spanish, starts_with("correct")) %>%
  gather(key = "measure", value = "score", starts_with("correct")) %>%
  mutate(measure = factor(gsub("_","",measure)))

study2_adjacentorswitch_whole$age.years.c <- scale(study2_adjacentorswitch_whole$age, scale = FALSE)[,1]

study2_adjacentorswitch <- study2_adjacentorswitch_whole %>%
  group_by(subj, age, age.years.c, gender, measure) %>%
  summarise(score = mean(score)) %>% ungroup()

study2_AdjSwitchModel <- summary(lme4::glmer(
  score ~ age.years.c * measure + (1|subj),
  data=study2_adjacentorswitch_whole, family = "binomial"))

study2_accuracyloess <- bind_cols(lapply(
  study2_adjacentorswitch %>% split(.$measure),
  function(x){
    as_data_frame(predict(loess(formula = score ~ age, data = x, span = 0.75), 
            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
      rename_all(funs(paste0(first(x$measure),"_",.)))
    })) %>%
  mutate(predicted_age = predicted_range) %>%
  gather(key = "measure", value = "score", starts_with("correct")) %>%
  separate("measure", into = c("measure", "stat"), sep = "_") %>%
  spread(stat, score)
  
# study2_accuracy_plot <- ggplot(study2_accuracyloess,
#   aes(x = predicted_age, y = fit, color = measure, group = measure)) + 
#   geom_line(size = 1) + geom_point(size = 4) +
#   geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = measure, fill = measure), alpha = .2, colour=NA) +
#   scale_color_viridis(discrete=TRUE, option = "D") +
#   scale_fill_viridis(discrete=TRUE, option = "D")


study2_EucDist <- study2_data %>%
  group_by(subj, age, gender) %>%
  filter(response_order %in% c("first", "only")) %>%
  mutate(wrong_euc_dist = ifelse(correct == 1, NA, wrong_euc_dist)) %>%
  filter(!is.na(wrong_euc_dist))

study2_EucDist$age.years.c <- scale(study2_EucDist$age, scale = FALSE)[,1]

# study2_EucDist_plot <- ggplot(study2_EucDist, aes(y = wrong_euc_dist, x = age)) +
#   geom_point(aes(color = spanish, group = spanish)) +
#   geom_smooth(method = "loess") +
#   scale_colour_manual(values = graph_colors)

study2_EucDist_model <- lme4::lmer(
  wrong_euc_dist ~ age.years.c + (1|subj),
  data=study2_EucDist, REML=FALSE)

study2_EucDist_pchisq <- pchisq(deviance(lm(wrong_euc_dist ~ 1, data=study2_EucDist)) - deviance(study2_EucDist_model), 78, lower.tail=FALSE)

```

<!-- The relationship between word entropy and language switching brings up the possibility that children may be using alternative strategies when they lack knowledge of the proper color term mapping.  -->

Another possible reason to use Spanish would be if children fail to recall the proper SK color term but do know the proper mapping in the Spanish. They may also choose to respond with a same-language but adjacent color term )such as "joshin" for a *panshin*-colored chip). In our next analysis, we aggregate across chips and examine the pattern of responses, categorizing them as same-language, adjacent, and different-language. 



```{r study2_termnature}

study2_allresponsetypes <- study2_data %>%
  filter(response_order %in% c("first", "only", "last")) %>%
  separate_rows(nature, sep = "-") %>%
  mutate(nature = trimws(ifelse(is.na(nature), "Other", nature))) %>%
  group_by(subj, age, response_order, nature) %>%
  summarise(n = n()) %>%
  group_by(subj, age, response_order) %>%
  mutate(prop = n/sum(n)) %>% ungroup() %>%
  complete(nesting(subj, age, response_order), nature, fill = list(n = 0, prop = 0.0))

study2_allresponsetypes_whole <- study2_allresponsetypes %>%
  select(-n) %>%
  spread(nature, prop)

study2_allresponsetypes_whole$age.years.c <- scale(study2_allresponsetypes_whole$age, scale = FALSE)[,1]
study2_allresponsetypes_whole$Y <- as.matrix(study2_allresponsetypes_whole[, c("BCT", "AHCT", "NBCT", "Other", "SLT")])


study2_allresponsetypes_model <- multinom(Y ~ age.years.c + response_order,
                 data=study2_allresponsetypes_whole)


study2_responsedev <- deviance(multinom(Y~1, data=study2_allresponsetypes_whole)) - deviance(multinom(Y ~ age.years.c , data=study2_allresponsetypes_whole))

study2_allresponseloess <- bind_rows(lapply(
  study2_allresponsetypes %>% split(list(.$response_order, .$nature)),
  function(x){
    if (nrow(x) > 0) {
    as_data_frame(predict(loess(formula = prop ~ age, data = x, span = 0.75),
            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
      mutate(predicted_age = predicted_range,
             nature = case_when(
               first(x$nature) == "AHCT" ~ "Ad-hoc color term",
               first(x$nature) == "BCT" ~ "Basic color term",
               first(x$nature) == "NBCT" ~ "Other color term",
               first(x$nature) == "SLT" ~ "Saturation/lightness term",
               TRUE ~ as.character(first(x$nature))),
             response_order = first(x$response_order))
    }}))

# study2_allresponseloess_plot <- ggplot(study2_allresponseloess,
#   aes(x = predicted_age, y = fit, color = nature, group = nature)) +
#   facet_grid(. ~ response_order) +
#   geom_line(size = 1) + geom_point(size = 4) +
#   geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = nature, fill = nature), alpha = .2, colour=NA) +
#   scale_color_viridis(discrete=TRUE, option = "D") +
#   scale_fill_viridis(discrete=TRUE, option = "D")

  
```


```{r study3_termnature}

study3_allresponsetypes <- study3_data %>%
  filter(response_order %in% c("first", "only", "last")) %>%
  separate_rows(nature, sep = "-") %>%
  mutate(nature = trimws(ifelse(is.na(nature), "Other", nature))) %>%
  group_by(subj, age, response_order, nature) %>%
  summarise(n = n()) %>%
  group_by(subj, age, response_order) %>%
  mutate(prop = n/sum(n)) %>% ungroup() %>%
  complete(nesting(subj, age, response_order), nature, fill = list(n = 0, prop = 0.0))


study3_allresponsetypes_whole <- study3_allresponsetypes %>%
  select(-n) %>%
  spread(nature, prop)

study3_allresponsetypes_whole$age.years.c <- scale(study3_allresponsetypes_whole$age, scale = FALSE)[,1]
study3_allresponsetypes_whole$Y <- as.matrix(study3_allresponsetypes_whole[, c("BCT", "AHCT", "NBCT", "Other", "SLT")])

study3_allresponsetypes_model <- multinom(Y ~ age.years.c + response_order,
                 data=study3_allresponsetypes_whole)

study3_allresponsetypes_dev <- deviance(multinom(Y~1, data=study3_allresponsetypes_whole)) - deviance(study3_allresponsetypes_model)


study3_allresponseloess <- bind_rows(lapply(
  study3_allresponsetypes %>% split(list(.$response_order, .$nature)),
  function(x){
    if (nrow(x) > 0) {
    as_data_frame(predict(loess(formula = prop ~ age, data = x, span = 0.75), 
            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
      mutate(predicted_age = predicted_range, 
             nature = case_when(
               first(x$nature) == "AHCT" ~ "Ad-hoc color term",
               first(x$nature) == "BCT" ~ "Basic color term",
               first(x$nature) == "NBCT" ~ "Other color term",
               first(x$nature) == "SLT" ~ "Saturation/lightness term",
               TRUE ~ as.character(first(x$nature))),
             response_order = first(x$response_order))
    }}))
  
# study3_allresponseloess_plot <- ggplot(study3_allresponseloess,
#   aes(x = predicted_age, y = fit, color = nature, group = nature)) + 
#   facet_grid(. ~ response_order) +
#   geom_line(size = 1) + geom_point(size = 4) +
#   geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = nature, fill = nature), alpha = .2, colour=NA) +
#   scale_color_viridis(discrete=TRUE, option = "D") +
#   scale_fill_viridis(discrete=TRUE, option = "D")
  
```

```{r study3_overextensions}

study3_adjacentorswitch_whole <- study3_data %>%
  group_by(subj, age, gender) %>%
  filter(response_order %in% c("first", "only")) %>%
  mutate(correct_nearby = correct + ifelse(is.na(wrong_adjacent), 0, wrong_adjacent),
         correct_all = ifelse(correct + correct_either + correct_nearby > 0, 1, 0)) %>%
  select(subj, age, gender, starts_with("correct")) %>%
  gather(key = "measure", value = "score", starts_with("correct")) %>%
  mutate(measure = factor(gsub("_","",measure)))

study3_adjacentorswitch_whole$age.years.c <- scale(study3_adjacentorswitch_whole$age, scale = FALSE)[,1]

study3_adjacentorswitch <- study3_adjacentorswitch_whole %>%
  group_by(subj, age, age.years.c, gender, measure) %>%
  summarise(score = mean(score)) %>% ungroup()

study3_AdjSwitchModel <- summary(lme4::glmer(
  score ~ age.years.c + measure + (1|subj),
  data=study3_adjacentorswitch_whole, family = "binomial"))

study3_accuracyloess <- bind_cols(lapply(
  study3_adjacentorswitch %>% split(.$measure),
  function(x){
    as_data_frame(predict(loess(formula = score ~ age, data = x, span = 0.75), 
            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
      rename_all(funs(paste0(first(x$measure),"_",.)))
    })) %>%
  mutate(predicted_age = predicted_range) %>%
  gather(key = "measure", value = "score", starts_with("correct")) %>%
  separate("measure", into = c("measure", "stat"), sep = "_") %>%
  spread(stat, score)
  
# study3_accuracy_plot <- ggplot(study3_accuracyloess,
#   aes(x = predicted_age, y = fit, color = measure, group = measure)) + 
#   geom_line(size = 1) + geom_point(size = 4) +
#   geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = measure, fill = measure), alpha = .2, colour=NA) +
#   scale_color_viridis(discrete=TRUE, option = "D") +
#   scale_fill_viridis(discrete=TRUE, option = "D")

study3_EucDist <- study3_data %>%
  group_by(subj, age, gender) %>%
  filter(response_order %in% c("first", "only")) %>%
  mutate(wrong_euc_dist = ifelse(correct == 1, NA, wrong_euc_dist)) %>%
  filter(!is.na(wrong_euc_dist))

study3_EucDist$age.years.c <- scale(study3_EucDist$age, scale = FALSE)[,1]

# study3_EucDist_plot <- ggplot(study3_EucDist, aes(y = wrong_euc_dist, x = age)) +
#   geom_point(aes(color = spanish, group = spanish)) +
#   geom_smooth(method = "loess") +
#   scale_colour_manual(values = graph_colors)

study3_EucDist_model <- lme4::lmer(
  wrong_euc_dist ~ age.years.c + (1|subj),
  data=study3_EucDist, REML=FALSE)

study3_EucDist_pchisq <- pchisq(deviance(lm(wrong_euc_dist ~ 1, data=study3_EucDist)) - deviance(study3_EucDist_model), 78, lower.tail=FALSE)


```

```{r crossstudy_accuracystrats}
bind_rows(
  list("Experiment 2" = study2_allresponseloess,
       "Experiment 3" = study3_allresponseloess), .id = "study") %>%
  mutate(response_order = tools::toTitleCase(as.character(response_order))) %>%
  ggplot(aes(x = predicted_age, y = fit, color = nature, group = nature)) +
  facet_grid(response_order ~ study) +
  geom_line() +
  geom_point() +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit,
                  group = nature, fill = nature), alpha = .2, colour=NA) +
  labs(x = "Age", y = "Proportion of Responses", fill = "Nature of Response", color = "Nature of Response") +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D") +
  coord_cartesian(ylim = c(0,1))
```

```{r, include = FALSE}
studies2n3_accuracy_plot <- ggplot(
  bind_rows(list("Experiment 2" = study2_accuracyloess, "Experiment 3" = study3_accuracyloess), .id = "study") %>%
    mutate(measure = case_when(
      measure == "correct" ~ "Original",
      measure == "correcteither" ~ "Either Language",
      measure == "correctnearby" ~ "Adjacent Color",
      measure == "correctall" ~ "Either + Adjacent")) %>%
    mutate(measure = factor(measure, levels = c("Original", "Either Language", "Adjacent Color", "Either + Adjacent"))),
  aes(x = predicted_age, y = fit, color = measure, group = measure)) +
  facet_grid(. ~ study) +
  geom_line(size = 1) + geom_point(size = 4) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = measure, fill = measure), alpha = .2, colour=NA) +
  labs(x = "Age", y = "Proportion of Responses", fill = "Accuracy", color = "Accuracy") +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D") +
  coord_cartesian(ylim = c(0,1))

# print(studies2n3_allresponseloess_plot)

# print(studies2n3_accuracy_plot)

```

Using a mixed-effects model, we found a significant improvement in accuracy scores when we allowed different-language but corresponding responses (*p* `r with(study2_AdjSwitchModel, ifelse(coefficients["measurecorrecteither", "Pr(>|z|)"] < 0.001, "< 0.001", paste0("= ", round(coefficients["measurecorrecteither", "Pr(>|z|)"], 3))))`) but no significant change when allowing for same-language but adjacent responses (*p* `r with(study2_AdjSwitchModel, ifelse(coefficients["measurecorrectnearby", "Pr(>|z|)"] < 0.001, "< 0.001", paste0("= ", round(coefficients["measurecorrectnearby", "Pr(>|z|)"], 3))))`).

**Comprehension.**



```{r study2_FLTypes, include = FALSE}

study2_FLTypes <- study2_data %>%
  filter(response_order %in% c("first", "last")) %>%
  select(subj, age, prompt, response_order, nature) %>%
  spread(response_order, nature) %>%
  mutate(first_BCT = ifelse(first == "BCT", 1, 0),
         first_AHCT = ifelse(first == "AHCT", 1, 0),
         last_BCT = ifelse(last == "BCT", 1, 0),
         last_AHCT = ifelse(last == "AHCT", 1, 0)) %>%
  group_by(subj, age) %>%
  summarise_at(vars(first_BCT:last_AHCT), mean, na.rm = TRUE)

# study2_FLTypes_plot <- ggplot(study2_FLTypes %>% 
#        gather(key = "measure", value = "prop", first_BCT:last_AHCT),
#        aes(y = prop, x = age, colour = measure, group = measure)) +
#   geom_smooth(method = "loess")

```


# Experiment 3

Noting the level of bilingualism in Experiment 2, we designed Experient 3 as its complement. In Experiment  3, we tested children entirely in Spanish with a set of chips representing prototypical colors for the Spanish color system.

```{r study3_participants}

study3_data <- feather::read_feather("../intermediates/processed_data.csv") %>%
  filter(language == "Spanish", task == "Production") %>%
  mutate(shipibo_response = ifelse(response %in% shipibo_terms, 1, 0)) 

study3_participants <- study3_data %>%
  group_by(subj) %>%
  filter(row_number() == 1) %>% ungroup() %>%
  summarise(min_age = min(age_interval),
            max_age = max(age_interval),
            num_56 = sum(age_interval %in% 5:6),
            num_78 = sum(age_interval %in% 7:8),
            num_910 = sum(age_interval %in% 9:10),
            num_1112 = sum(age_interval %in% 11:12),
            num_benajema = sum(location == 'bena jema'),
            num_bilingue = sum(location == 'bilingue'),
            num_ccnn = sum(location == 'c.c.n.n. (bahuanisho)'),
            num_boys = sum(gender == "masculino"),
            num_total = n())

study3_agesex <- study3_data %>%
  group_by(subj) %>%
  filter(row_number() == 1) %>% group_by(age_interval) %>%
  summarise(num_total = n(),
            num_boys = sum(gender == 'masculino'),
            mean_age = round(mean(age), 1),
            sd_age = round(sd(age), 2)) %>% ungroup() %>%
  mutate(freq_total = round(100*num_total/sum(num_total)),
         freq_boys = round(100*num_boys/num_total)) %>%
  mutate(`Age Group` = paste0(age_interval, "-years-old"),
         `Age (SD)` = paste0(mean_age, " (", sd_age, ")"),
         `N` = paste0(num_total, " (", freq_total, "%)"),
         `Male` = paste0(num_boys, " (", freq_boys, "%)")) %>%
  select(`Age Group`, `Age (SD)`, `N`, `Male`)

study3_table1 <- datatable(study3_agesex, rownames = FALSE)

```

### Participants

Our protocol received ethical approval from Pontificia Universidad Católica del Perú’s Institutional Review Board. Children were recruited in a SK neighborhood of Yarinacocha (Bena Jema) as well as in Bawanisho. As before, children were recruited either through their parents or through the local school. When recruited at school, consent for participation was collected from both the teachers and the parents; otherwise, only consent from the parents was collected. Data were collected from a total of 46 children (16 boys) who were between the ages of 5 and 11 years old.

### Materials

Even though participants of Experiment 1 were instructed to give color terms in SK, some Spanish color terms were provided (this was especially true of young adult participants). Based on these data and on previous studies of Spanish color systems, we singled out 11 color chips that were prototypical instances of prominent Peruvian Spanish color terms. These color chips were grey (WCS n°46), pink (WCS n°65), orange (WCS n°121), green (WCS n°234), red (WCS n°245), brown (WCS n°266), white (WCS n°274), blue (WCS n°291), yellow (WCS n°297), black (WCS n°312) and purple (WCS n°325). (To visualize the hue of these chips, see Appendix 1.) These color chips were exactly the same as those used in Experiment 1; the only difference was that while 330 chips were used in Experiment 1, only 11 of them were used in Experiment 3. It is worth noting that six chips were shared between Experiment 2 and Experiment 3.

### Procedure

Since SK children are not very fluent in Spanish, the production and comprehension tasks were both conducted in SK, and Spanish was only used for color terms (i.e., Spanish color terms were embedded in SK sentences). In both tasks, children were seating in front of the experimenter; a table (on which the color chips were displayed) was standing between them. As in Experiment 2, the production task was always performed before the comprehension task.

Production task. The procedure was the same as that of Experiment 2. Children were first introduced to the whole procedure and the general goal of the study. It was specified that they would be expected to provide color terms in Spanish (and not in SK). Children were then asked: "what is the color of this chip?". When children provided SK color terms, the experimenter would write down their response but further ask: "what is the name of this color in Spanish?". When children were replying "I don’t know" to this prompt, the experimenter would not ask further questions and would move forward to the next color chip. As a result, responses of some children include only non-basic Spanish color terms or SK color terms. In total, we collected production data for 11 color chips. For each chip, the data include either one response (when children provided a Spanish basic color term in the first trial) or two or three responses (when children’s initial responses were either non-basic and/or in SK).

Comprehension task. The procedure was similar to that of the comprehension task of Experiment 2. The 11 color chips of the production task were simultaneously displayed in front of the children. The experimenter would then ask: "Can you give me the ___ chip?" (where "___" stands for a color term). In total, the comprehension of 11 Spanish color terms was tested.

The choice of these terms was based on previous studies examining Spanish color terms as well as on Experiment 1 (as we have seen, SK adults sometimes resorted to Spanish color terms to name the color chips). The 11 terms used as prompts included: blanco ("white"), verde ("green"), rojo ("red"), amarillo ("yellow"), azul ("blue"), negro ("black"), naranja ("orange"), gris ("grey"), morado ("purple"), marrón ("brown"), and rosa ("pink").
Since each color term was instantiated by only one color chip, no term required the special procedure that was followed in Experiment 2 for yankon, joshin and pei/xo. 


## Results and Discussion.

```{r study3_spanishentropy}
study3_switchperc <- study3_data %>%
  group_by(subj) %>%
  summarise(shipibo_response = 100*mean(shipibo_response)) %>%
  ungroup() %>%
  summarise(mean_switch = mean(shipibo_response),
            sd_switch = sd(shipibo_response),
            min_switch = min(shipibo_response),
            max_switch = max(shipibo_response)) %>%
  mutate_all(funs(paste0(round(.),'%')))

study3_overallscore <- study3_data %>%
  group_by(subj, age) %>%
  summarise(correct = mean(correct, na.rm = T))

study3_overallscore_test <- cor.test(~ correct + age, data = study3_overallscore)

study3_overallswitch <- study3_data %>%
  group_by(subj, age) %>%
  summarise(shipibo_response = mean(shipibo_response))

study3_overallswitch_test <- cor.test(~ shipibo_response + age, data = study3_overallswitch)

study3_responsebyentropy <- study3_data %>%
  filter(spanish %in% c("Blanco", "Negro", "Verde", "Amarillo", "Rojo") & 
           (task == "Comprehension" | response_order %in% c("first", "only"))) %>%
  left_join(study1_entropy %>% mutate(`Chip ID` = as.character(`Chip ID`)),
            by = c("prompt" = "Chip ID"))

study3_responsebyentropy$age.years.c <- scale(study3_responsebyentropy$age, scale = FALSE)[,1]

study3_SwitchEnt_model <- summary(lme4::glmer(
  shipibo_response ~ age.years.c * Entropy + (1|subj),
  data=study3_responsebyentropy, family = "binomial"))

```

Similar to Experiment 2, over a quarter of all responses (*M* = `r study3_switchperc$mean_switch`, *SD* = `r study3_switchperc$sd_switch`) were given in another language (Shipibo in this case). There was significant variation in language-switching with some children completing the entire task in Spanish while others responded to upwards of `r study3_switchperc$max_switch` of trials in Shipibo. Similar to Experiment 2, there was no significant correlation between age and label accuracy (*p* `r with(study3_overallscore_test, ifelse(p.value < 0.001, "< 0.001", paste0("= ", round(p.value, 3))))`) or between age and language-switching (*p* `r with(study3_overallswitch_test, ifelse(p.value < 0.001, "< 0.001", paste0("= ", round(p.value, 3))))`). Still, we found that participants tended to respond in Shipibo when presented with items that had low entropy among SK adults during Experiment 1 (p `r with(study3_SwitchEnt_model, ifelse(coefficients["Entropy", "Pr(>|z|)"] < 0.001, "< 0.001", paste0("= ", round(coefficients["Entropy", "Pr(>|z|)"], 3))))`). This suggests that participants across Studies 2 and 3 preferred to respond in Shipibo when presented with a high-consensus chip and in Spanish when shown a low-consensus chip.

**Overextensions.**


Similar to Experiment 2, we adopted alternative scoring to accommodate language-switching from Spanish to Shipibo-Konibo and same-language adjacent responses. Using a mixed-effects model, we did not find that age explained a significant amount of the variation seen in accuracy (*p* `r with(study3_AdjSwitchModel, ifelse(coefficients["age.years.c", "Pr(>|z|)"] < 0.001, "< 0.001", paste0("= ", round(coefficients["age.years.c", "Pr(>|z|)"], 3))))`), in concordance with earlier analyses. However, we did find that participants made use of both mapping strategies of either providing different-language but corresponding responses (*p* `r with(study3_AdjSwitchModel, ifelse(coefficients["measurecorrecteither", "Pr(>|z|)"] < 0.001, "< 0.001", paste0("= ", round(coefficients["measurecorrecteither", "Pr(>|z|)"], 3))))`) or same-language but adjacent responses (*p* `r with(study3_AdjSwitchModel, ifelse(coefficients["measurecorrectnearby", "Pr(>|z|)"] < 0.001, "< 0.001", paste0("= ", round(coefficients["measurecorrectnearby", "Pr(>|z|)"], 3))))`). Between Studies 2 and 3, we find frequent use of language switching but only Experiment 3 shows significant use of same-language but adjacent terms as well. This discrepancy, along with the lack of an age correlation, can be due to foreign language exposure. Children may be exposed to Spanish at a young age but do not receive any formal Spanish education until later in adolescence. With a limited knowledge of Spanish color terms, children may spontaneously provide Spanish color terms during the Shipibo-language Experiment 2 but may struggle to succeed during Spanish-language Experiment 3. This suggests that children may rely on either strategy to communicate a color label to the best of their knowledge set.



```{r study3_FLTypes, include = FALSE}

study3_FLTypes <- study3_data %>% ungroup() %>%
  filter(response_order %in% c("first", "last")) %>%
  select(subj, age, prompt, response_order, nature) %>%
  spread(response_order, nature) %>%
  mutate(first_BCT = ifelse(first == "BCT", 1, 0),
         first_AHCT = ifelse(first == "AHCT", 1, 0),
         last_BCT = ifelse(last == "BCT", 1, 0),
         last_AHCT = ifelse(last == "AHCT", 1, 0)) %>%
  group_by(subj, age) %>%
  summarise_at(vars(first_BCT:last_AHCT), mean, na.rm = TRUE)

ggplot(study3_FLTypes %>% gather(key = "measure", value = "prop", first_BCT:last_AHCT),
       aes(y = prop, x = age, colour = measure, group = measure)) +
  geom_smooth(method = "loess")


study3_FLTypes$age.years.c <- scale(study3_FLTypes$age, scale = FALSE)[,1]
study3_FLTypes$Y <- as.matrix(study3_FLTypes[, c("first_BCT", "first_AHCT", "last_BCT", "last_AHCT")])

study3_allresponsetypes_model <- multinom(Y ~ age.years.c + response_order,
                 data=study3_allresponsetypes_whole)

study3_allresponsetypes_dev <- deviance(multinom(Y~1, data=study3_allresponsetypes_whole)) - deviance(study3_allresponsetypes_model)

pchisq(study3_allresponsetypes_dev, 39+39-2, lower.tail = T)

```


**Comparisons between Studies 2 & 3. - Unfinished plots**

```{r crossstudy_adhoc, include = FALSE}

studies2n3_data <- feather::read_feather("../intermediates/processed_data.csv") %>%
  filter(task == "Production") %>%
  mutate(language = case_when(
      language == "Shipibo" ~ "Experiment 2",
      language == "Spanish" ~ "Experiment 3"
    ),
    spanish_response = ifelse(response %in% spanish_terms, 1, 0),
    shipibo_response = ifelse(response %in% shipibo_terms, 1, 0)) %>%
  rename(study = language)

studies2n3_firstresponse_whole <- studies2n3_data %>% ungroup() %>%
  filter(response_order %in% c("first", "only")) %>%
  select(study, subj, age, prompt, nature, spanish_response, shipibo_response) %>%
  left_join(mutate(study1_entropy, `Chip ID` = as.character(`Chip ID`)), 
            by = c("prompt" = "Chip ID"))

studies2n3_firstresponse_whole$age.years.c <- scale(studies2n3_firstresponse_whole$age, scale = FALSE)[,1]
studies2n3_firstresponse_whole$nature <- relevel(studies2n3_firstresponse_whole$nature, ref = "BCT")

study2_naturebyentropy_model <- multinom(nature ~ age.years.c + Entropy + (1:subj),
                 data=filter(studies2n3_firstresponse_whole, study == "Experiment 2"))
study3_naturebyentropy_model <- multinom(nature ~ age.years.c + Entropy + (1:subj),
                 data=filter(studies2n3_firstresponse_whole, study == "Experiment 3"))

studies2n3_naturebyentropy_plot <- ggplot(studies2n3_firstresponse_whole %>%
         mutate(BCT = as.numeric(grepl("BCT", nature)),
                AHCT = as.numeric(grepl("AHCT", nature))) %>%
         group_by(study, prompt, Entropy) %>%
         mutate_at(vars(BCT, AHCT), funs(ifelse(is.na(.), 0, .))) %>%
         summarise_at(vars(BCT, AHCT), mean, na.rm = TRUE) %>%
         gather(key = "nature", value = "prop", BCT, AHCT) %>%
         mutate(nature = case_when(
                  nature == "BCT" ~ "Basic Color Term",
                  nature == "AHCT" ~ "Ad-hoc color term"),
                nature = relevel(factor(nature), ref = "Basic Color Term")), 
       aes(x = Entropy, y = prop, group = nature, colour = nature, fill = nature)) +
  facet_grid(. ~ study) +
  geom_smooth(method = "lm") +
  coord_cartesian(ylim = c(0,1)) +
    labs(x = "Labelling Entropy of Chip", y = "Proportion of Responses", color = "Nature of Response", fill = "Nature of Response") +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D") 

studies2n3_firstresponse <- studies2n3_firstresponse_whole %>%
  group_by(study, subj, age, nature, spanish_response, shipibo_response) %>%
  summarise(n = n()) %>% 
  group_by(study, subj, age) %>%
  mutate(prop = n/sum(n),
         response_language = case_when(
           spanish_response == 1 ~ "spanish",
           shipibo_response == 1 ~ "shipibo",
           TRUE ~ NA_character_
         )) %>% ungroup()


studies2n3_firstresploess <- bind_rows(lapply(
  studies2n3_firstresponse %>% 
    filter(nature %in% c("AHCT", "BCT")) %>%
    split(list(.$study, .$nature, .$response_language)),
  function(x){
    if (nrow(x) > 6) {
    as_data_frame(predict(loess(formula = prop ~ age, data = x, span = 0.75), 
            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
      mutate(predicted_age = predicted_range,
             response_language = tools::toTitleCase(first(x$response_language)),
             study = first(x$study),
             nature = case_when(
               first(x$nature) == "AHCT" ~ "Ad-hoc color term",
               first(x$nature) == "BCT" ~ "Basic color term",
               TRUE ~ "Other"))
    }}))

studies2n3_firstresploess_plot <- ggplot(studies2n3_firstresploess,
  aes(x = predicted_age, y = fit, color = response_language, group = response_language)) + 
  facet_grid(nature ~ study) +
  geom_line(size = 1) + geom_point(size = 4) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = response_language, fill = response_language), alpha = .2, colour=NA) +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D") +
  labs(x = "Age", y = "Proportion of Responses", colour = "Language of Response", fill = "Language of Response") +
  coord_cartesian(ylim = c(0,1))
print(studies2n3_naturebyentropy_plot)

print(studies2n3_firstresploess_plot)

```





# General Discussion

TO BE PASTED FROM GOOGLE DOC

\newpage

# References

```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
