```{r study2_termnature}

study2_allresponsetypes <- study2_data %>%
  filter(response_order %in% c("first", "only", "last")) %>%
  separate_rows(nature, sep = "-") %>%
  mutate(nature = trimws(ifelse(is.na(nature), "Other", nature))) %>%
  group_by(subj, age, response_order, nature) %>%
  summarise(n = n()) %>%
  group_by(subj, age, response_order) %>%
  mutate(prop = n/sum(n)) %>% ungroup() %>%
  complete(nesting(subj, age, response_order), nature, fill = list(n = 0, prop = 0.0))

study2_allresponsetypes_whole <- study2_allresponsetypes %>%
  select(-n) %>%
  spread(nature, prop)

study2_allresponsetypes_whole$age.years.c <- scale(study2_allresponsetypes_whole$age, scale = FALSE)[,1]
study2_allresponsetypes_whole$Y <- as.matrix(study2_allresponsetypes_whole[, c("BCT", "AHCT", "NBCT", "Other", "SLT")])


study2_allresponsetypes_model <- multinom(Y ~ age.years.c + response_order,
                                          data=study2_allresponsetypes_whole)


study2_responsedev <- deviance(multinom(Y~1, data=study2_allresponsetypes_whole)) - deviance(multinom(Y ~ age.years.c , data=study2_allresponsetypes_whole))

study2_allresponseloess <- bind_rows(lapply(
  study2_allresponsetypes %>% split(list(.$response_order, .$nature)),
  function(x){
    if (nrow(x) > 0) {
      as_data_frame(predict(loess(formula = prop ~ age, data = x, span = 0.75),
                            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
        mutate(predicted_age = predicted_range,
               nature = case_when(
                 first(x$nature) == "AHCT" ~ "Ad-hoc color term",
                 first(x$nature) == "BCT" ~ "Basic color term",
                 first(x$nature) == "NBCT" ~ "Other color term",
                 first(x$nature) == "SLT" ~ "Saturation/lightness term",
                 TRUE ~ as.character(first(x$nature))),
               response_order = first(x$response_order))
    }}))

# study2_allresponseloess_plot <- ggplot(study2_allresponseloess,
#   aes(x = predicted_age, y = fit, color = nature, group = nature)) +
#   facet_grid(. ~ response_order) +
#   geom_line(size = 1) + geom_point(size = 4) +
#   geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = nature, fill = nature), alpha = .2, colour=NA) +
#   scale_color_viridis(discrete=TRUE, option = "D") +
#   scale_fill_viridis(discrete=TRUE, option = "D")


```






```{r study2_FLTypes, include = FALSE}

study2_FLTypes <- study2_data %>%
  filter(response_order %in% c("first", "last")) %>%
  select(subj, age, prompt, response_order, nature) %>%
  spread(response_order, nature) %>%
  mutate(first_BCT = ifelse(first == "BCT", 1, 0),
         first_AHCT = ifelse(first == "AHCT", 1, 0),
         last_BCT = ifelse(last == "BCT", 1, 0),
         last_AHCT = ifelse(last == "AHCT", 1, 0)) %>%
  group_by(subj, age) %>%
  summarise_at(vars(first_BCT:last_AHCT), mean, na.rm = TRUE)

# study2_FLTypes_plot <- ggplot(study2_FLTypes %>% 
#        gather(key = "measure", value = "prop", first_BCT:last_AHCT),
#        aes(y = prop, x = age, colour = measure, group = measure)) +
#   geom_smooth(method = "loess")

```





study3_accuracy_plot <- ggplot(study3_accuracyloess,
                               aes(x = predicted_age, y = fit, color = measure, group = measure)) +
  geom_line(size = 1) + geom_point(size = 4) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = measure, fill = measure), alpha = .2, colour=NA) +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D")

study3_EucDist <- study3_data %>%
  group_by(subj, age, gender) %>%
  filter(response_order %in% c("first", "only")) %>%
  mutate(wrong_euc_dist = ifelse(correct == 1, NA, wrong_euc_dist)) %>%
  filter(!is.na(wrong_euc_dist))

study3_EucDist$age.years.c <- scale(study3_EucDist$age, scale = FALSE)[,1]

# study3_EucDist_plot <- ggplot(study3_EucDist, aes(y = wrong_euc_dist, x = age)) +
#   geom_point(aes(color = spanish, group = spanish)) +
#   geom_smooth(method = "loess") +
#   scale_colour_manual(values = graph_colors)

study3_EucDist_model <- lme4::lmer(
  wrong_euc_dist ~ age.years.c + (1|subj),
  data=study3_EucDist, REML=FALSE)

study3_EucDist_pchisq <- pchisq(deviance(lm(wrong_euc_dist ~ 1, data=study3_EucDist)) - deviance(study3_EucDist_model), 78, lower.tail=FALSE)




```{r}
study2_EucDist <- study2_data %>%
  group_by(subj, age, gender) %>%
  filter(response_order %in% c("first", "only")) %>%
  mutate(wrong_euc_dist = ifelse(correct == 1, NA, wrong_euc_dist)) %>%
  filter(!is.na(wrong_euc_dist))

study2_EucDist$age.years.c <- scale(study2_EucDist$age, scale = FALSE)[,1]

# study2_EucDist_plot <- ggplot(study2_EucDist, aes(y = wrong_euc_dist, x = age)) +
#   geom_point(aes(color = spanish, group = spanish)) +
#   geom_smooth(method = "loess") +
#   scale_colour_manual(values = graph_colors)

study2_EucDist_model <- lme4::lmer(
  wrong_euc_dist ~ age.years.c + (1|subj),
  data=study2_EucDist, REML=FALSE)

study2_EucDist_pchisq <- pchisq(deviance(lm(wrong_euc_dist ~ 1, data=study2_EucDist)) - deviance(study2_EucDist_model), 78, lower.tail=FALSE)
```




```{r study3_FLTypes, include = FALSE}

study3_FLTypes <- study3_data %>% ungroup() %>%
  filter(response_order %in% c("first", "last")) %>%
  select(subj, age, prompt, response_order, nature) %>%
  spread(response_order, nature) %>%
  mutate(first_BCT = ifelse(first == "BCT", 1, 0),
         first_AHCT = ifelse(first == "AHCT", 1, 0),
         last_BCT = ifelse(last == "BCT", 1, 0),
         last_AHCT = ifelse(last == "AHCT", 1, 0)) %>%
  group_by(subj, age) %>%
  summarise_at(vars(first_BCT:last_AHCT), mean, na.rm = TRUE)

ggplot(study3_FLTypes %>% gather(key = "measure", value = "prop", first_BCT:last_AHCT),
       aes(y = prop, x = age, colour = measure, group = measure)) +
  geom_smooth(method = "loess")


study3_FLTypes$age.years.c <- scale(study3_FLTypes$age, scale = FALSE)[,1]
study3_FLTypes$Y <- as.matrix(study3_FLTypes[, c("first_BCT", "first_AHCT", "last_BCT", "last_AHCT")])

study3_allresponsetypes_model <- multinom(Y ~ age.years.c + response_order,
                                          data=study3_allresponsetypes_whole)

study3_allresponsetypes_dev <- deviance(multinom(Y~1, data=study3_allresponsetypes_whole)) - deviance(study3_allresponsetypes_model)

pchisq(study3_allresponsetypes_dev, 39+39-2, lower.tail = T)

```



```{r crossstudy_accuracystrats}
bind_rows(
  list("Experiment 2" = study2_allresponseloess,
       "Experiment 3" = study3_allresponseloess), .id = "study") %>%
  mutate(response_order = tools::toTitleCase(as.character(response_order))) %>%
  ggplot(aes(x = predicted_age, y = fit, color = nature, group = nature)) +
  facet_grid(response_order ~ study) +
  geom_line() +
  geom_point() +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit,
                  group = nature, fill = nature), alpha = .2, colour=NA) +
  labs(x = "Age", y = "Proportion of Responses", fill = "Nature of Response", color = "Nature of Response") +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D") +
  coord_cartesian(ylim = c(0,1))
```




**Comparisons between Studies 2 & 3. - Unfinished plots**
  
  ```{r crossstudy_adhoc, include = FALSE}

studies2n3_data <- feather::read_feather("../intermediates/processed_data.csv") %>%
  filter(task == "Production") %>%
  mutate(language = case_when(
    language == "Shipibo" ~ "Experiment 2",
    language == "Spanish" ~ "Experiment 3"
  ),
  spanish_response = ifelse(response %in% spanish_terms, 1, 0),
  shipibo_response = ifelse(response %in% shipibo_terms, 1, 0)) %>%
  rename(study = language)

studies2n3_firstresponse_whole <- studies2n3_data %>% ungroup() %>%
  filter(response_order %in% c("first", "only")) %>%
  select(study, subj, age, prompt, nature, spanish_response, shipibo_response) %>%
  left_join(mutate(study1_entropy, `Chip ID` = as.character(`Chip ID`)), 
            by = c("prompt" = "Chip ID"))

studies2n3_firstresponse_whole$age.years.c <- scale(studies2n3_firstresponse_whole$age, scale = FALSE)[,1]
studies2n3_firstresponse_whole$nature <- relevel(studies2n3_firstresponse_whole$nature, ref = "BCT")

study2_naturebyentropy_model <- multinom(nature ~ age.years.c + Entropy + (1:subj),
                                         data=filter(studies2n3_firstresponse_whole, study == "Experiment 2"))
study3_naturebyentropy_model <- multinom(nature ~ age.years.c + Entropy + (1:subj),
                                         data=filter(studies2n3_firstresponse_whole, study == "Experiment 3"))

studies2n3_naturebyentropy_plot <- ggplot(studies2n3_firstresponse_whole %>%
                                            mutate(BCT = as.numeric(grepl("BCT", nature)),
                                                   AHCT = as.numeric(grepl("AHCT", nature))) %>%
                                            group_by(study, prompt, Entropy) %>%
                                            mutate_at(vars(BCT, AHCT), funs(ifelse(is.na(.), 0, .))) %>%
                                            summarise_at(vars(BCT, AHCT), mean, na.rm = TRUE) %>%
                                            gather(key = "nature", value = "prop", BCT, AHCT) %>%
                                            mutate(nature = case_when(
                                              nature == "BCT" ~ "Basic Color Term",
                                              nature == "AHCT" ~ "Ad-hoc color term"),
                                              nature = relevel(factor(nature), ref = "Basic Color Term")), 
                                          aes(x = Entropy, y = prop, group = nature, colour = nature, fill = nature)) +
  facet_grid(. ~ study) +
  geom_smooth(method = "lm") +
  coord_cartesian(ylim = c(0,1)) +
  labs(x = "Labelling Entropy of Chip", y = "Proportion of Responses", color = "Nature of Response", fill = "Nature of Response") +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D") 

studies2n3_firstresponse <- studies2n3_firstresponse_whole %>%
  group_by(study, subj, age, nature, spanish_response, shipibo_response) %>%
  summarise(n = n()) %>% 
  group_by(study, subj, age) %>%
  mutate(prop = n/sum(n),
         response_language = case_when(
           spanish_response == 1 ~ "spanish",
           shipibo_response == 1 ~ "shipibo",
           TRUE ~ NA_character_
         )) %>% ungroup()


studies2n3_firstresploess <- bind_rows(lapply(
  studies2n3_firstresponse %>% 
    filter(nature %in% c("AHCT", "BCT")) %>%
    split(list(.$study, .$nature, .$response_language)),
  function(x){
    if (nrow(x) > 6) {
      as_data_frame(predict(loess(formula = prop ~ age, data = x, span = 0.75), 
                            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
        mutate(predicted_age = predicted_range,
               response_language = tools::toTitleCase(first(x$response_language)),
               study = first(x$study),
               nature = case_when(
                 first(x$nature) == "AHCT" ~ "Ad-hoc color term",
                 first(x$nature) == "BCT" ~ "Basic color term",
                 TRUE ~ "Other"))
    }}))

studies2n3_firstresploess_plot <- ggplot(studies2n3_firstresploess,
                                         aes(x = predicted_age, y = fit, color = response_language, group = response_language)) + 
  facet_grid(nature ~ study) +
  geom_line(size = 1) + geom_point(size = 4) +
  geom_ribbon(aes(ymin = fit - se.fit, ymax = fit + se.fit, group = response_language, fill = response_language), alpha = .2, colour=NA) +
  scale_color_viridis(discrete=TRUE, option = "D") +
  scale_fill_viridis(discrete=TRUE, option = "D") +
  labs(x = "Age", y = "Proportion of Responses", colour = "Language of Response", fill = "Language of Response") +
  coord_cartesian(ylim = c(0,1))
print(studies2n3_naturebyentropy_plot)

print(studies2n3_firstresploess_plot)

```





```{r study3_termnature}

study3_allresponsetypes <- study3_data %>%
  filter(response_order %in% c("first", "only", "last")) %>%
  separate_rows(nature, sep = "-") %>%
  mutate(nature = trimws(ifelse(is.na(nature), "Other", nature))) %>%
  group_by(subj, age, response_order, nature) %>%
  summarise(n = n()) %>%
  group_by(subj, age, response_order) %>%
  mutate(prop = n/sum(n)) %>% ungroup() %>%
  complete(nesting(subj, age, response_order), nature, fill = list(n = 0, prop = 0.0))


study3_allresponsetypes_whole <- study3_allresponsetypes %>%
  select(-n) %>%
  spread(nature, prop)

study3_allresponsetypes_whole$age.years.c <- scale(study3_allresponsetypes_whole$age, scale = FALSE)[,1]
study3_allresponsetypes_whole$Y <- as.matrix(study3_allresponsetypes_whole[, c("BCT", "AHCT", "NBCT", "Other", "SLT")])


study3_allresponseloess <- bind_rows(lapply(
  study3_allresponsetypes %>% split(list(.$response_order, .$nature)),
  function(x){
    if (nrow(x) > 0) {
      as_data_frame(predict(loess(formula = prop ~ age, data = x, span = 0.75), 
                            newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
        mutate(predicted_age = predicted_range, 
               nature = case_when(
                 first(x$nature) == "AHCT" ~ "Ad-hoc color term",
                 first(x$nature) == "BCT" ~ "Basic color term",
                 first(x$nature) == "NBCT" ~ "Other color term",
                 first(x$nature) == "SLT" ~ "Saturation/lightness term",
                 TRUE ~ as.character(first(x$nature))),
               response_order = first(x$response_order))
    }}))
```

study2_accuracyloess <- bind_cols(lapply(
  study2_adjacentorswitch %>% 
    split(.$measure),
  function(x){
    as_data_frame(predict(loess(formula = score ~ age, data = x, span = 0.75), 
                          newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
      rename_all(funs(paste0(first(x$measure),"_",.)))
  })) %>%
  mutate(predicted_age = predicted_range) %>%
  gather(key = "measure", value = "score", starts_with("correct")) %>%
  separate("measure", into = c("measure", "stat"), sep = "_") %>%
  spread(stat, score)

study3_accuracyloess <- bind_cols(lapply(
  study3_adjacentorswitch %>% split(.$measure),
  function(x){
    as_data_frame(predict(loess(formula = score ~ age, data = x, span = 0.75), 
                          newdata = predicted_range, se = TRUE)[c("fit", "se.fit")]) %>%
      rename_all(funs(paste0(first(x$measure),"_",.)))
  })) %>%
  mutate(predicted_age = predicted_range) %>%
  gather(key = "measure", value = "score", starts_with("correct")) %>%
  separate("measure", into = c("measure", "stat"), sep = "_") %>%
  spread(stat, score)

